Blame and Coercion: Together Again for the Second Time

### 2019-10-17 paper.tex
revise bg:cast calc

\begin{figure}
	Syntax
	\[
	\begin{array}{rclr}
	\stxrule{T}{types}{
		\star \mid{}
		P
	}
	\stxrule{P}{pre-types}{
		\POOunit \mid
		\POOfun{T_1}{T_2} \mid
		\POOprod{T_1}{T_2} \mid
		\POOsum{T_1}{T_2}
	}
	\stxrule{I}{injectable types (\lazyD)}{
		P
	}
	\stxrule{I}{injectable types (\lazyUD)}{
		\POOunit \mid
		\POOfun{\star}{\star} \mid
		\star \times \star \mid
		\star + \star
	}
	\stxrule{e}{terms}{
		\eOOvar{x} \mid{}
		\eOOsole{} \mid{}
		\eOOlam{T_1}{T_2}{x}{e} \mid
		\eOOapp{e_1}{e_2}
	}
	\stxrulecont{
		\eOOcons{e_1}{e_2} \mid
		\eOOcar{e} \mid
		\eOOcdr{e}
	}
	\stxrulecont{
		\eOOinl{e} \mid
		\eOOinr{e} \mid
		\eOOcase{e_1}{e_2}{e_3}
	}
	\stxrulecont{
		\eOOcast{e}{c} \mid
		\eOOblame{l}
	}
	\stxrule{v}{values}{
		\vOOfun{\rho}{x}{e} \mid
		\vOOtt{} \mid
		\vOOcons{v_1}{v_2} \mid
		\vOOinl{v} \mid
		\vOOinr{v} \mid		
		\vOOcast{v}{c}
	}
	\stxrule{c}{casts}{
		\cOOcast{T_1}{l}{T_2}
	}
	\stxrule{o}{observations}{
		\oOOinj \mid
		\oOOsole \mid
		\oOOfun \mid
		\oOOcons \mid
		\oOOinl \mid
		\oOOinr \mid
		\oOOblame{l}
	}
	\end{array}
	\]

@@

\figref{fig:blame-static} defines the syntax of cast calculus and its
static semantics. We extend the syntax in \citet{siek2009exploring} with sum 
types and product types. Besides, we annotate function codomain explicitly 
because we need the codomain when evaluating cast calculus with hypercoercions.
\todo[inline]{Which previous definitions are you refering to?  What
  are the little changes? -Jeremy
\\ --- \\ Fixed -KC}
\todo[inline]{The term typing rule for casts is broken. -Jeremy
\\ --- \\ Fixed -KC}
\todo[inline]{Why do you annotate the codomain of lambda abstractions
	explicitly? -Jeremy
	\\---\\
	Because I apply $ id(\cdot) $ to the codomain in the HC CEK. -KC}
Let $ S,T $ range over types. A type is either the dynamic type $ \star $
(a.k.a. $ \mathtt{Dyn} $, $ \mathbb{?} $, or $ \mathtt{Unknown} $), 
or a pre-type. 
Let $ P,Q $ range over pre-types. Every pre-type is a type with a type 
constructor at the top. For simplicity, we only have one base type, $ \POOunit 
$. 
Other pre-types are functions, products, and sums.

%$ S \sim T $ reads ``$ S $ and $ T $ are consistent''.
%Two types are consistent if one of them is $ \star $, or they have the same 
%top-most type constructor and the corresponding sub-parts are consistent. 
%The intuition of $ S \sim T $ is that $ S $ and $ T $ have no conflict type 
%information. Consistency is reflexive and symmetric, but not transitive.

%One might be tempted to conclude that inconsistency is the root of all runtime 
%type errors (blames). In the setting of gradual typing, however, runtime 
%type-checking is usually gradual as well. 
$ S \smile T $ reads ``$ S $ and $ T $ are shallowly-consistent''. Two types 
are shallowly consistent if one of them is $ \star $, or they have the same 
type constructor at the top. Shallow-inconsistency is the root of all blame in 
lazy cast strategies -- casting a value to a shallowly inconsistent type leads 
to a blame. Shallow-consistency is reflexive, symmetric, but 
not transitive.

Let $ e $ range over terms, including all terms from $ \mathtt{STLC+} $ and, 
in addition, casts and blame. Unlike $ \mathtt{STLC+} $, we annotate the 
codomain of lambda abstractions explicitly.
The end of \figref{fig:blame-static} shows the typing rules for the
additional terms.

>>>>

\begin{figure}
	Syntax
	\[
	\begin{array}{rclr}
	\stxrule{T}{types}{
		\star \mid{}
		P
	}
	\stxrule{P}{pre-types}{
		\POOunit \mid
		\POOfun{T_1}{T_2} \mid
		\POOprod{T_1}{T_2} \mid
		\POOsum{T_1}{T_2}
	}
	\stxrule{c}{casts}{
		\cOOcast{T_1}{l}{T_2}
	}
	\stxrule{e}{terms}{
		\eOOvar{x} \mid{}
		\eOOsole{} \mid{}
		\eOOlam{T_1}{T_2}{x}{e} \mid
		\eOOapp{e_1}{e_2}
	}
	\stxrulecont{
		\eOOcons{e_1}{e_2} \mid
		\eOOcar{e} \mid
		\eOOcdr{e}
	}
	\stxrulecont{
		\eOOinl{e} \mid
		\eOOinr{e} \mid
		\eOOcase{e_1}{e_2}{e_3}
	}
	\stxrulecont{
		\eOOcast{e}{c} \mid
		\eOOblame{l}
	}
	\stxrule{I}{injectable types (\lazyD)}{
		P
	}
	\stxrule{I}{injectable types (\lazyUD)}{
		\POOunit \mid
		\POOfun{\star}{\star} \mid
		\star \times \star \mid
		\star + \star
	}
	\stxrule{v}{values}{
		\vOOfun{\rho}{x}{e} \mid
		\vOOtt{} \mid
		\vOOcons{v_1}{v_2} \mid
		\vOOinl{v} \mid
		\vOOinr{v} \mid		
		\vOOcast{v}{c}
	}
	\stxrule{o}{observations}{
		\oOOinj \mid
		\oOOsole \mid
		\oOOfun \mid
		\oOOcons \mid
		\oOOinl \mid
		\oOOinr \mid
		\oOOblame{l}
	}
	\end{array}
	\]

@@

\figref{fig:blame-static} defines the syntax of cast calculus and its
static semantics. We extend the syntax in \citet{siek2009exploring} with sum 
types and product types. Besides, we annotate function codomain explicitly 
because we need the codomain when evaluating cast calculus with hypercoercions.
\todo[inline]{Which previous definitions are you refering to?  What
  are the little changes? -Jeremy
\\ --- \\ Fixed -KC}
\todo[inline]{The term typing rule for casts is broken. -Jeremy
\\ --- \\ Fixed -KC}
\todo[inline]{Why do you annotate the codomain of lambda abstractions
	explicitly? -Jeremy
	\\---\\
	Because I apply $ id(\cdot) $ to the codomain in the HC CEK. -KC}
Let $ T $ range over types. A type is either the dynamic type $ \star $
(a.k.a. $ \mathtt{Dyn} $, $ \mathbb{?} $, or $ \mathtt{Unknown} $), 
or a pre-type. 
Let $ P $ range over pre-types. Every pre-type is a type with a type 
constructor at the top. For simplicity, we only have one base type, $ \POOunit 
$. 
Other pre-types are functions, products, and sums.

$ T_1 \sim T_2 $ reads ``$ T_1 $ and $ T_2 $ are consistent''.
The intuition of $ T_1 \sim T $ is that $ T_1 $ and $ T_2 $ have no conflict 
type information. Two types are consistent if one of them is $ \star $, or they 
have the same top-most type constructor and the corresponding sub-parts are 
consistent. Consistency is reflexive and symmetric, but not transitive.

$ T_1 \smile T_2 $ reads ``$ T_1 $ and $ T_2 $ are shallowly-consistent''. Two 
types are shallowly consistent if one of them is $ \star $, or they have the 
same type constructor at the top. Shallow-inconsistency is the root of all 
blame in lazy cast strategies -- casting a value to a shallowly inconsistent 
type leads to a blame. Shallow-consistency is reflexive, symmetric, but 
not transitive.

Let $ e $ range over terms. Most terms are standard. We annotate the codomain 
of lambda abstractions explicitly because we need it when evaluating cast 
calculus with hypercoercion. We have two additional terms: cast expressions and 
blame expressions. Cast expressions insert runtime type checks. And blame 
expressions raise an error. The end of \figref{fig:blame-static} shows the 
typing rules for the additional terms.
